Efficient and Accurate Modeling of Flow through Sedimentary Structures with MODFLOW 6

### 2023-10-23 latex/article.tex
provost edits
various edits aimed at fleshing things out, flow, and consistent style

\textbf{Article impact statement:} {\color{red} AMP: Need impact statement}

\begin{abstract}
\noindent {\color{red} AMP: compose Abstract near the end}
\end{abstract}

@@

The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2022}. It attempts to simulate uniform flow in an inclined permeable hydrogeologic layer (``aquifer'') embedded in less permeable material (the ``domain''). The aquifer is of uniform width and hydraulic conductivity and is inclined relative to the horizontal.

To simulate flow along an aquifer inclined at angle $\theta$ ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using a formula that corresponds to a uniform, unit head gradient:
\begin{equation}
\label{eqn:head_analyt_along}
h = - x \cos \theta - z \sin \theta.
\end{equation}

\noindent where $h$ is head in meters and $x$ and $z$ are the horizontal and vertical model coordinates, respectively, in meters. The hydraulic conductivity of the aquifer is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d along the aquifer. In the base case, the conductivity of the domain is $10^{-6}$ m/d to effectively isolate the aquifer hydraulically.

Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically offset grid of MODFLOW 6 DISU (unstructured) type. The cell geometry in these two discretization methods is identical, but the cell connectivity is different. The DIS grid uses layered connectivity, and the DISU grid includes the additional connections between cells in adjacent grid layers needed to realize full connectivity. The DISU grid is constructed by starting with a DIS grid, converting it to an equivalent DISU grid, and then implementing full connectivity. The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the aquifer, 3 layers in the domain above the aquifer, and 3 layers in the domain below the aquifer.

The above description represents the base case test problem. However, the results also include simulations designed to investigate the effects of varying the gridding resolution, the dip angle of the aquifer, the conductivity contrast between the aquifer and the domain, and the anisotropy of the conductivity tensor.

@@

The test problem allows the numerical solution for head and fluxes to be compared with an analytical solution. This section presents results for the base case scenario when comparing gridding strategy (i.e., layered connectivity versus full connectivity) and flow solution formulation (i.e., standard conductance based formulation versus XT3D option) to test the hypothesis that a grid with full connectivity produces a more accurate solution for dipping hydrogeologic layers than a grid with only layered connectivity. The base case scenario is then modified to further investigate other aspects including grid resolution, dip angle, conductivity contrast and anisotropy.

Results for flux magnitude, flux direction and volumetric flow through the aquifer are used to assess the accuracy of the numerical representations of the test problem. Note that the terms ``flux''  and ``specific discharge'' are used interchangeably. Reported fluxes for each model are for the central model cell and are therefore uninfluenced by the model boundary or the interface between the aquifer and the domain. Volumetric flow through the aquifer is therefore also examined as it encompasses fluxes through the entire aquifer. This paper uses the sum of the flows reported for the CHD cells along the left end of the aquifer as the volumetric flow.

\subsection{Gridding Strategy and Flow Formulation Comparison}

@@

Through working on this test problem, we have encountered some nuances in working with specific discharges calculated using MODFLOW. There are two main MODFLOW output files pertaining to fluxes. The first is an array of face flows which specifies the total flow across each cell face (``FLOW-JA-FACE''). The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS''), and is calculated using distance-weighted averages of the cell face flows. Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the aquifer edge are ``dampened'' due to averaging of the face flows from the cell top and bottom. The result is a ``staircase'' effect whereby these edge cells have a reduced flux magnitude and ``flattened'' flux directions, as shown in Figure 3c of \cite{bardot2022}. The plots in Figure \ref{fig:fig2} display corrected cell-averaged specific discharges which exclude the near-zero face flows along the aquifer boundary, and therefore represents the actual simulated flux magnitude and direction within the cell.

@@

\begin{figure}[p!]
	\begin{center}
	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
	\caption{Numerical results for the test problem using base case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation. The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange). The right panel shows the face flows (red arrows).}
	\label{fig:fig2}
	\end{center}
\end{figure}

\subsection{Grid Resolution}

@@

\begin{figure}
	\begin{center}
	\includegraphics[scale=0.9]{../figures/fig3paper.png}
	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the hydrogeologic layer. We immediately notice the errors layered-connectivity grids incur (blue line), and how using a full-connectivity grid reproduces the analytical solution from two grid layers used to discretise the hydrogeologic layer (orange).}
	\label{fig:fig3}
	\end{center}
\end{figure}

The base scenario arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}. Layered-connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one grid layer per hydrogeological unit. However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine (Figure \ref{fig:fig3}). We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution.

Flux magnitude error for layered-connectivity grids peaked at around 10\% but improved and tended to zero with refinement. Flux direction for full-connectivity grids is clearly a major issue, being reported as horizontal for one grid layer per aquifer width, and only improved to 12$^{\circ}$ for 9 grid layers per aquifer width. Flow through the aquifer peaked at 13\% error but decreased to only a 4\% error for 9 grid layers per aquifer width.

@@

\subsection{Dip angle and conductivity contrast}

The base case scenario uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between aquifer and domain of $1:10^{-6}$. Here, we investigate the behavior of the flow solution for a wide range of aquifer inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80 $^{\circ}$. Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the aquifer. Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.

Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}). Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30 $^{\circ}$ when the solution starts peeling away from the analytical. Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios. The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65 $^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps. The full-connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65 $^{\circ}$, with the exception of a slight decrease in flux direction after 45 $^{\circ}$.

@@

\subsection{Anisotropy}

The base case and subsequent scenarios have assumed isotropic conductivity. However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors {\color{red} (AMP: throughout, or only in the aquifer?)}, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000. We also include the isotropic scenario (ratio of 1). Results are presented in Figure \ref{fig:fig5}. 
Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full-connectivity grids reproduce the analytical solution (orange line). These results verify, once again, our proposed method of using full-connectivity grids to model dipping anisotropic hydrogeologic layers.

@@

\cite{bardot2022} revisited the capability of MODFLOW to accurately simulate groundwater flow through sedimentary structures.  They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows for rectilinear grid overlay approaches in which locations of sedimentary structures are mapped onto a relatively fine model grid.  They also concluded that the XT3D multi-point flux approximation did not perform as well as anticipated for vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.  They hypothesized that the inherent connectivity in DIS and DISV layered grids in MODFLOW 6 was inadequate for allowing the XT3D multi-point flux approximation to work as intended.

In this paper, we confirm that the layered connectivity implemented in DIS and DISV model grids is inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flux approximation is used.  However, when additional cross connections are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.  Importantly, this accuracy improvement requires that at least two grid layers are used to represent the steeply dipping sedimentary structure.  The primary conclusion of this paper, therefore, is that that the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through sedimentary structures.  Prior to this work, the capability to efficiently and accurately model flow through sedimentary structures was generally restricted to finite element simulators.

This paper presents a solution for accurately representing steeply dipping flow structures groundwater utilising MODFLOW 6. \citep{bardot2022} demonstrated that the flux magnitude in highly dipping scenarios was erroneous when utilising layered-connectivity grids.  {\color{red} CDL: Also applicable for MODFLOW-USG DISU grids.} More significantly, the direction of fluxes were inaccurate, neglecting the vertical component of the flux. The issues observed in the existing study of \citep{bardot2022} partially related to the implementation of only within-grid-layer horizontal connections, limiting the model capacity to simulate the vertical component of the flow. The present study developed an approach which utilised full-connectivity grids \cite{modflow6gwf} to overcome the perceived problems. Through comparison of the previous layered-connectivity and proposed full-connectivity approach to the analytical solution presented in \citep{bardot2022}, we demonstrate that the use of full-connectivity grids (in conjunction with the XT3D formulation) is an accurate method to simulate flow through dipping hydrogeologic layers. However, accurate representation of flows requires a minimum of two grid layers per hydrogeologic unit to allow for vertical connections to induce flows.

@@

2) The concepts in this paper have been easiest to discuss in terms of grid layers, but they're not limited to layered grids.

3) (Random thought) The use of full-connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the aquifer geometry preformed better in isotropic grids. If you force the normal flux to be the same at the interface of the domain and aquifer, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the aquifer correctly?  Would figure \ref{fig:fig2} show a correction for this if the anisotropic scenario was tested?

@@

\section{Acknowledgments}
We would like to acknowledge....

>>>>

\textbf{Article impact statement: Groundwater flow through dipping sedimentary structures can be accurately simulated using the XT3D multi-point flow formulation in MODFLOW 6.}

\begin{abstract}
Groundwater flow through dipping hydrogeologic layers is often modeled with MODFLOW by specifying top and bottom elevations of model layers to coincide with top and bottom elevations of hydrogeologic layers.  With this approach, adjacent cells within a model layer are vertically offset from one another.  While the standard formulation in MODFLOW does not rigorously account for these offsets in the underlying flow calculations, results are generally expected to be accurate provided dip angles do not exceed $10^{\circ}$.  To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D multi-point flow formulation was introduced in MODFLOW 6.  The XT3D flow formulation is designed to account for geometric irregularities in the grid, including vertical offsets between laterally adjacent cells, and to provide accurate results for both isotropic and anisotropic groundwater flow.  \cite{bardot2022} showed, however, that neither the standard formulation nor the XT3D formulation produced the correct flow direction and magnitude in simple tests involving flow through a steeply dipping, permeable hydrogeologic layer.  In this paper, we explain that the inability of XT3D to give accurate results in the steeply dipping test problem of \cite{bardot2022} was caused by the limited cell connectivity that is inherent in the most commonly used discretization packages in MODFLOW 6. The DIS and DISV discretization packages are based on layered connectivity in which cells have lateral connections only with cells in the same grid layer.  We demonstrate that layered lateral connectivity is insufficient to allow accurate simulation of flow through a steeply dipping hydrogeologic layer on a vertically offset grid.  When lateral connections between cells in adjacent grid layers are included using the fully unstructured (DISU) grid type, XT3D is able to calculate the flow magnitude and direction accurately for both isotropic flow and for anisotropy oriented along the permeable hydrogeologic layer.  To realize this improved accuracy, the permeable hydrogeologic layer must be discretized vertically using at least two model layers. {\color{red} (AMP: Given the idealized nature of the test problem, how general can we be with our conclusions? I pulled it back here relative to what you had originally, CDL, but left a slightly stronger statement in the Conclusions, at least for now.})  Our findings suggest that, given appropriate discretization and cell connectivity, the combination of vertically offset grids with the XT3D flow formulation in MODFLOW 6 can provide a computationally efficient and accurate method for simulating isotropic and anisotropic flow through steeply dipping sedimentary structures, a capability that was previously reserved to finite-element-based simulators. {\color{red} (AMP: Any Pal and Edwards-type MPFA stuff that's relevant to that final phrase? See also Conclusions)}
\end{abstract}

@@

The test problem is patterned after the cross-sectional (transect) benchmark of \cite{bardot2022}. It attempts to simulate uniform flow in an inclined permeable hydrogeologic layer (``aquifer'') embedded in less permeable material (the ``domain'').

To simulate flow along an aquifer of uniform thickness inclined at angle $\theta$ relative to the horizontal ($30^{\circ}$ for the base case), heads are specified at the centers of cells along the perimeter of the model using a formula that corresponds to a uniform, unit head gradient:
\begin{equation}
\label{eqn:head_analyt_along}
h = - x \cos \theta - z \sin \theta.
\end{equation}

\noindent where $h$ is head in meters and $x$ and $z$ are the horizontal and vertical model coordinates, respectively, in meters. The hydraulic conductivity of the aquifer is set to 1 m/d (meter per day) so that the analytical flow solution is a groundwater flux of 1 m/d along the aquifer. In the base case, the conductivity of the domain surrounding the aquifer is set to $10^{-6}$ m/d to effectively isolate the aquifer hydraulically.

Two methods of discretization are used to simulate the test problem: (1) vertically offset grid of the MODFLOW 6 DIS (structured) type, and (2) vertically offset grid of MODFLOW 6 DISU (unstructured) type. The cell geometry in these two discretization methods is identical, but the cell connectivity is different. The DIS grid uses layered connectivity, and the DISU grid includes the additional connections between cells in adjacent grid layers needed to realize full connectivity. The DISU grid is constructed by starting with a DIS grid, converting it to an equivalent DISU grid, and then implementing full connectivity. The base case uses grids consisting of 11 columns and 9 layers of cells, with 3 layers within the aquifer, 3 layers in the domain above the aquifer, and 3 layers in the domain below the aquifer. {\color{red} (AMP: Need dimensions of aquifer and domain.)}

The above description represents the base-case test problem. However, the results also include simulations designed to investigate the effects of varying the gridding resolution, the dip angle of the aquifer, the conductivity contrast between the aquifer and the domain, and the anisotropy of the conductivity tensor.

@@

{\color{red} (AMP: Did some editing of this section for flow and consistent style, but I'll have another go once we're satisfied that all the pieces are basically in place. We're really close, if not there already.)}

To demonstrate the effect of cell connectivity on the accuracy of simulated flow through a steeply dipping hydrogeologic layer (aquifer), results for the base case described in the previous section are compared for simulations that use either layered or full connectivity, and either the standard or the XT3D flow formulation. The base case is then modified to further investigate the effects of grid resolution, aquifer dip angle, conductivity contrast between the aquifer and the surrounding domain, and anisotropic hydraulic conductivity.

Plots of specific discharge (groundwater flux) at cell centers and flows across cell-cell interfaces (face flows) calculated by MODFLOW 6 are used to illustrate the accruacy of the simulated flows. The magnitude and direction of specific discharge at the cell in the center of the aquifer, at which boundary effects are presumably minimized, and the total volumetric flow rate entering the aquifer through the left-hand boundary, which consists of constant-head (CHD) cells, are also reported as ``representative'' values of flow. {\color{red} (AMP: Heads are also plotted. Do we want to also report, say, maximum error in head?)}

Specific discharge at cell centers is calculated in a postprocessing step using distance-weighted interpolation of face flows calculated by MODFLOW 6 and reported in the binary budget output file. For the purpose of illustrating and evaluating the results of this idealized benchmark, specific discharge in cells within the aquifer is calculated using only face flows between cells within the aquifer. Likewise, specific discharge in cells within the surrounding domain is calculated using only face flows between cells within the domain. Specific discharge values reported in the binary budget output file by MODFLOW 6 are different than the values reported here for cells that are adjacent to the boundary between the aquifer and the domain. This is because MODFLOW interpolates specific discharge in a cell using all of the face flows for that cell, including flows across cell faces on the aquifer-domain boundary. Those flows represent an ``average'' of the different flows on either side of the aquifer-domain boundary. Therefore, specific discharge calculated in a cell adjacent to the aquifer-domain boundary is affected by the different flow rate on the other side of the boundary. While this is an unavoidable consequence of vertically offset discretization in practical applications, and not necessarily a problem, users should be aware of these effects when interpreting the results of MODFLOW 6 dispersive-transport and MODPATH particle-tracking simulations.

\subsection{Effects of Cell Connectivity and Flow Formulation}

@@

The results shown in the top four panels of Figure \ref{fig:fig2} {\color{red} (AMP: Label the figure panels a through h so we can refer to them by letter?)}, which are based on a vertically offset grid with layered connectivity, are consistent with the findings of \cite{bardot2022}. Despite boundary heads designed to induce specific discharge of 1 m/d along the $30^{\circ}$ incline of the aquifer, the simulated flow within the aquifer is predominantly horizontal, and the magnitude of the specific discharge at the center of the aquifer and the total flow entering the aquifer are both overestimated. The XT3D flow formulation is unable to compensate for the inadequate cell connectivity, which does not allow flow between laterally adjacent cells in different grid layers.

The results shown in the bottom four panels of Figure \ref{fig:fig2}, which are based on a vertically offset grid with full connectivity, show significantly improved results, even for the standard flow formulation. Using the XT3D flow formulation provides the most accurate results, as expected.

%Through working on this test problem, we have encountered some nuances in working with specific discharges calculated using MODFLOW. There are two main MODFLOW output files pertaining to fluxes. The first is an array of face flows which specifies the total flow across each cell face (``FLOW-JA-FACE''). The second is the cell-averaged specific discharge for each cell in its directional components ($q_{x}$, $q_{y}$, $q_{z}$) (``DATA-SPDIS''), and is calculated using distance-weighted averages of the cell face flows. Therefore, the specific discharge magnitude and direction normally reported by MODFLOW for cells within and along the aquifer edge are ``dampened'' due to averaging of the face flows from the cell top and bottom. The result is a ``staircase'' effect whereby these edge cells have a reduced flux magnitude and ``flattened'' flux directions, as shown in Figure 3c of \cite{bardot2022}. The plots in Figure \ref{fig:fig2} display corrected cell-averaged specific discharges which exclude the near-zero face flows along the aquifer boundary, and therefore represents the actual simulated flux magnitude and direction within the cell. {\color{red} (AMP: An edited version of this now appears near the beginning of the Results and Discussion section)}

@@

\begin{figure}[p!]
	\begin{center}
	\includegraphics[scale=0.8]{../figures/fig2_paper.png}
	\caption{Numerical results for the test problem using base-case settings for layered-connectivity and full-connectivity grids, with the standard conductance based formulation as well as the XT3D formulation. The left panel of each scenario shows the calculated specific discharge at each cell centre (black arrows) with head contours (orange). The right panel shows the face flows (red arrows). {\color{red} (AMP: Change ``Vertically Offset'' to ``Layered Connectivity'' and ``Vertically Staggered'' to ``Full Connectivity'' in the script that generates the figure)} {\color{red} (AMP: Magnify tiny vectors in the domain [by the conductivity contrast, say] so they can be seen? Face flows along the aquifer/domain boundary are tricky because they're ``in between'' in terms of magnitude.)}}
	\label{fig:fig2}
	\end{center}
\end{figure}

\subsection{Effects of Grid Resolution}

@@

\begin{figure}
	\begin{center}
	\includegraphics[scale=0.9]{../figures/fig3paper.png}
	\caption{Graphs indicating how increasing the number of grid layers within a single hydrogeologic layer (in this case, the aquifer) affects (a) flux magnitude of middle cell, (b) flux direction of middle cell and (c) volumetric flow through the hydrogeologic layer. We immediately notice the errors layered-connectivity grids incur (blue line), and how using a full-connectivity grid reproduces the analytical solution from two grid layers used to discretise the hydrogeologic layer (orange). {\color{red} (AMP: Change ``Vertically Offset'' to ``Layered Connectivity'', ``Vertically Staggered'' to ``Full Connectivity'', and ``Model layers'' to ``Grid layers'' in the script that generates the figure)}}
	\label{fig:fig3}
	\end{center}
\end{figure}

The base case arbitrarily uses 3 grid layers to represent the aquifer, but it is important to address the question: \emph{How many grid layers are required to adequately represent a hydrogeological layer?}. Layered-connectivity grids, which have been used by groundwater modellers for decades (ref?) {\color{red} Jim to check, I think there is some refinement guidance for aquitards}, are almost always configured as one grid layer per hydrogeological unit. {\color{red} (AMP: Could maybe mention that it's generally well known you want multiple layers for accurate transport?)} However, although computationally efficient, we investigate the effect of altering resolution within the aquifer by using the base case and varying number of cells per aquifer width from one to nine {\color{red} (AMP: Explain how aquifer and domain dimensions vary with the discretization. And is that how we want them to vary?)} Figure \ref{fig:fig3}). {\color{red} (AMP: Are all results in this figure with XT3D? If so, would there be any benefit to showing the standard formulation results, as well?)} We compare modeled flux magnitude and direction in the center cell, as well as the volumetric flow against the analytical solution. {\color{red} (AMP: Should we also say something about the error in the head solution? Report the maximum head error in each case?)}

Flux magnitude error for layered-connectivity grids peaked at around 10\% but improved and tended to zero with refinement. Flux direction for layered-connectivity grids is clearly a major issue, being reported as horizontal for one grid layer per aquifer width, and only improved to 12$^{\circ}$ for 9 grid layers per aquifer width. Flow through the aquifer peaked at 13\% error but decreased to only a 4\% error for 9 grid layers per aquifer width.

@@

\subsection{Effects of Dip Angle and Conductivity Contrast}

The base case uses a dip angle of 30$^{\circ}$ and an extreme conductivity contrast between aquifer and domain of 1:$10^{-6}$. Here, we investigate the behavior of the flow solution for a wide range of aquifer inclines by systematically changing the dip angle by 2.5$^{\circ}$ from 0 to 80$^{\circ}$. Similarly, we examine the solution for multiple conductivity contrasts by changing the domain conductivity to 2, 5, 10 and 100 times less than the aquifer. {\color{red} (AMP: Dare we try making the domain more conductive than the ``aquifer''?)}  Simulated flux magnitude and direction of the centre cell for all combinations are plotted against the analytical solution in Figure \ref{fig:fig4}.

Traditional layered-connectivity grids without XT3D shows rapid diversion of flux magnitude and direction from the analytical solution as dip increases (Figure \ref{fig:fig4a}). Use of the XT3D option keeps the flux magnitude somewhat on-track until about 30$^{\circ}$ when the solution starts peeling away from the analytical. Flux direction is improved for the homogeneous scenario, but not for the heterogeneous scenarios. The full-connectivity grid on the other hand, even without XT3D, roughly follows the analytical solution for flux magnitude and direction until around 65$^{\circ}$ when it starts to ``fall-apart.'' There are also spikes at certain dip angles which are related to transitions between cell connections due to changing cell face overlaps. The full-connectivity grid with XT3D clearly boasts the best results with the flux magnitude and direction matching the analytical until 65$^{\circ}$, with the exception of a slight decrease in flux direction after 45$^{\circ}$. {\color{red} (AMP: Can explain the wheels coming off around 65 deg by the fact that with only 3 cells vertically in the aquifer, the aquifer then looks like a long, zigzag Tetris piece that's no more than one cell wide, either vertically or horizontally.  Finer vertical discretization improves things.)}

@@

\subsection{Effect of Anisotropy}

The base case and subsequent scenarios have assumed isotropic conductivity. However, sedimentary layers often exhibit anisotropic conductivity, and therefore we examine the gridding strategies using anisotropic and dipping conductivity tensors {\color{red} (AMP: in the domain and aquifer, or only in the aquifer?)}, with reduced conductivity perpendicular to the aquifer by 10, 100, 1000 and 10,000. We also include the isotropic scenario (ratio of 1). {\color{red} (AMP: Dare we try increased perpendicular conductivity?)} Results are presented in Figure \ref{fig:fig5}. {\color{red} (AMP: Are all results in this figure with XT3D? If so, would there be any benefit to showing the standard formulation results, as well?)} Layered-connectivity grids produce flux magnitude and flow results that significantly deviate from the analytical solution (blue line), whilst full-connectivity grids reproduce the analytical solution (orange line). These results verify, once again, our proposed method of using full-connectivity grids to model dipping anisotropic hydrogeologic layers.

@@

\cite{bardot2022} revisited the capability of MODFLOW to accurately simulate groundwater flow through sedimentary structures.  They concluded that the XT3D multi-point flux approximation in MODFLOW 6 significantly improves the accuracy of simulated flows when used with a grid-overlay approach in which locations of sedimentary structures, which can be steeply dipping, are mapped onto a relatively fine, rectilinear model grid.  They also concluded that the XT3D multi-point flux approximation does not perform as well as anticipated when simulating flow through a steeply dipping hydrogeologic layer using vertically offset grids, which are advantageous for significantly reducing the number of grid layers required for many problems.  They hypothesized that the inherently limited lateral connectivity between cells in DIS and DISV layered grids in MODFLOW 6 was responsible for the inability to obtain accurate flow results in their steeply dipping test problems using vertically offsets grids, with or without XT3D.

In this paper, we have confirmed that the layered lateral connectivity implemented in DIS and DISV model grids is inadequate for simulating groundwater flow through steeply dipping sedimentary structures, even when the XT3D multi-point flux approximation is used.  However, when connections between laterally adjacent cells in different grid layers are added to comprise a grid with full connectivity, there is a substantial increase in the accuracy of simulated flows.  The additional connectivity is implemented using the fully unstructured (DISU) grid type.  Importantly, this accuracy improvement requires that at least two grid layers be used to represent the steeply dipping sedimentary structure.  The primary conclusion of this paper is that, given appropriate discretization and cell connectivity, the XT3D multi-point flux approximation implemented in MODFLOW 6 can be combined with vertically offset grids to efficiently and accurately model flows through {\color{red} steeply dipping??} sedimentary structures.  Prior to this work, the capability to efficiently and accurately model flow through {\color{red} steeply dipping??} sedimentary structures was generally restricted to finite element simulators.

The importance of adequate connectivity has been explained and evaluated in this work specifically in terms of hydrogeologic layers that correspond to sedimentary structures. However, the impact of cell connectivity on the accuracy of simulated flows should be considered in any MODFLOW application that models flow through a heterogeneous hydraulic conductivity field using a grid with large vertical offsets between laterally adjacent cells, whether or not the grid has a layered structure. The full connectivity needed to improve accuracy can be implemented using the DISU grid types in both MODFLOW 6 and MODFLOW-USG.

@@

3) (Random thought) The use of full-connectivity grids improved the accuracy of the transect solutions; however, the results of \citep{bardot2022} using unstructured meshes that conformed to the aquifer geometry preformed better in isotropic grids. If you force the normal flux to be the same at the interface of the domain and aquifer, the analytical solution would suggest a uniform gradient either side, do we introduce an error if the faces don't align with the aquifer correctly?  Would figure \ref{fig:fig2} show a correction for this if the anisotropic scenario was tested?  {\color{red} (AMP: Is the underlying question whether using an unstructured mesh that conforms to the aquifer geometry [which MF6 can't do in transect] would improve the accuracy more than a full-connectivity vertically offset grid, assuming xt3d is used in both cases? Let's discuss on our next call.)}

@@

\section{Acknowledgments}
We would like to acknowledge.... {\color{red} that this simple paper has been in the works far too long. Kindly address complaints to the first author.}
